---
title: "Data Visualization"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(plotly)
library(viridis)
cleaned_data = read_csv("cleaned_data.csv")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)

theme_set (theme_minimal() +theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplots.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_disrete = scale_fill_viridis_d
```


## overweight percentage changes over year
```{r message = FALSE, echo = FALSE}
# add up each county's percentage for each year
county_year_data = 
  cleaned_data |>
  group_by (county, year) |>
  summarize(ave_overweight_percent = mean(percent_overweight)) 

ggplot(county_year_data, aes(x = year, y = ave_overweight_percent, group = county, color = county )) + 
  geom_line() + 
  geom_point() +
  labs(title = "Average Overweight Percentage Over Time",
       x = "Year",
       y = "Average Overweight Percentage") +
  theme_minimal() +
  theme(legend.position = "none")
  
```
<br>
interpretation: This plot looks messy, needs to think about it later

or only consider average overweight percentage over year?
```{r message = FALSE, echo = FALSE}
# add up each county's percentage for each year
county_year_data = 
  cleaned_data |>
  group_by (year) |>
  summarize(ave_overweight_percent = mean(percent_overweight)) 

ggplot(county_year_data, aes(x = year, y = ave_overweight_percent)) + 
  geom_line() + 
  geom_point() +
  labs(title = "Average Overweight Percentage Over Time",
       x = "Year",
       y = "Average Overweight Percentage") +
  theme_minimal() +
  theme(legend.position = "none")
```

## County level number of percent overweight/obese/healthy
```{r message = FALSE, echo = FALSE}
# calculate the average percentage of overweight/obsed/healthy for all years
county_ave_data = 
  cleaned_data |>
  group_by (county) |>
  summarize(ave_overweight_percent = mean(percent_overweight), 
            ave_obese_percent = mean(percent_obese), 
            ave_healthy_percent = mean(percent_healthy_weight)*100) 

plot_ly(county_ave_data, x = ~county, y = ~ave_overweight_percent, type = 'bar', color = ~county) |>
  layout(title = "Percentage Overweight by County",
         xaxis = list(title = "County"),
         yaxis = list(title = "Average Percentage Overweight"),
         showlegend = FALSE,
         xaxis = list(tickangle = -60))

plot_ly(county_ave_data, x = ~county, y = ~ave_obese_percent, type = 'bar', color = ~county) |>
  layout(title = "Percentage Obese by County",
         xaxis = list(title = "County"),
         yaxis = list(title = "Average Percentage Obese"),
         showlegend = FALSE,
         xaxis = list(tickangle = -60))

plot_ly(county_ave_data, x = ~county, y = ~ave_healthy_percent, type = 'bar', color = ~county) |>
  layout(title = "Percentage Healthy by County",
         xaxis = list(title = "County"),
         yaxis = list(title = "Average Percentage Healthy"),
         showlegend = FALSE,
         xaxis = list(tickangle = -60))
```
Interpretation: 


## elementary and middle school overweight across county
```{r message = FALSE, echo = FALSE}
element_middle_data = 
  cleaned_data |>
  group_by (county, grade_level) |>
  filter(grade_level == "ELEMENTARY" | grade_level == "MIDDLE/HIGH") |>
  summarize(ave_overweight_percent = mean(percent_overweight), 
            ave_obese_percent = mean(percent_obese), 
            ave_healthy_percent = mean(percent_healthy_weight)*100) 
```


```{r message = FALSE, echo = FALSE}
# display and compare elementary and middle/high overweight data  
ggplot(element_middle_data, aes(x = county, y = ave_overweight_percent, fill = grade_level)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Overweight Percentage by County and Grade Level",
       x = "County",
       y = "Overweight Percentage",
       fill = "Grade Level") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## sex and overweight
```{r message = FALSE, echo = FALSE}
sex_data = 
  cleaned_data |>
  group_by (sex, year) |>
  summarize(ave_overweight_percent = mean(percent_overweight), 
            ave_obese_percent = mean(percent_obese), 
            ave_healthy_percent = mean(percent_healthy_weight)*100) 
```

```{r message = FALSE, echo = FALSE}
# geom line to compare 
ggplot(sex_data, aes(x = year, y = ave_overweight_percent, color = sex)) +
  geom_point() +
  geom_line() +
  labs(title = "Average Overweight Percentage Over Years by Sex",
       x = "Year",
       y = "Average Overweight Percentage",
       color = "Sex") +
  theme_minimal()
```

## racial group and overweight
```{r message = FALSE, echo = FALSE}
# the distribution of racial group, first group by district zip code
race_dis_data = cleaned_data |>
  group_by (year, district) |>
  summarize (num_white = mean(num_white), 
             num_asian = mean(num_asian), 
             num_am_ind = mean(num_am_ind), 
             num_black = mean(num_black), 
             num_hisp = mean(num_hisp))

race_year_data = race_dis_data |>
  group_by (year) |>
  summarize(asian = sum(num_asian), 
            white = sum(num_white), 
            am_ind = sum(num_am_ind), 
            black = sum(num_black), 
            hisp = sum(num_hisp))

race_year_data = pivot_longer(race_year_data, cols = -year, names_to = "race", values_to = "number" )

```

```{r message = FALSE, echo = FALSE}
race_2015 = race_year_data |>
  filter(year == "2015")

plot_2015 <- plot_ly(race_2015, labels = ~race, values = ~ number, type = "pie", hole = 0.2, textinfo = 'number', textposition = 'inside', marker = list(colors = c("pink", "orange", "brown", "blue", "green"))) |>
  layout(title = "Distribution of Races in 2015")

race_2016 = race_year_data |>
  filter(year == "2016")

plot_2016 = plot_ly(race_2016, labels = ~race, values = ~ number, type = "pie", hole = 0.2, textinfo = 'number', textposition = 'inside', marker = list(colors = c("pink", "orange", "brown", "blue", "green"))) |>
  layout(title = "Distribution of Races in 2016")

race_2017 = race_year_data |>
  filter(year == "2017")

plot_2017 = plot_ly(race_2017, labels = ~race, values = ~ number, type = "pie", hole = 0.2, textinfo = 'number', textposition = 'inside', marker = list(colors = c("pink", "orange", "brown", "blue", "green"))) |>
  layout(title = "Distribution of Races in 2017")

race_2018 = race_year_data |>
  filter(year == "2018")

plot_2018 = plot_ly(race_2018, labels = ~race, values = ~ number, type = "pie", hole = 0.2, textinfo = 'number', textposition = 'inside', marker = list(colors = c("pink", "orange", "brown", "blue", "green"))) |>
  layout(title = "Distribution of Races in 2018")

race_2019 = race_year_data |>
  filter(year == "2019")

plot_2019 = plot_ly(race_2019, labels = ~race, values = ~ number, type = "pie", hole = 0.2, textinfo = 'number', textposition = 'inside', marker = list(colors = c("pink", "orange", "brown", "blue", "green"))) |>
  layout(title = "Distribution of Races in 2019")

plot_2015
plot_2016
plot_2017
plot_2018
plot_2019
```

county level racial group and county level overweight
```{r message = FALSE, echo = FALSE}
county_dis_data = cleaned_data |>
  group_by(year, county, district) |>
  summarize (percent_overweight = mean(percent_overweight), 
             num_white = mean(num_white), 
             num_asian = mean(num_asian), 
             num_am_ind = mean(num_am_ind), 
             num_black = mean(num_black), 
             num_hisp = mean(num_hisp))

county_year_data = county_dis_data |>
  group_by(year, county) |>
  summarize (percent_overweight = mean(percent_overweight), 
             num_white = sum(num_white), 
             num_asian = sum(num_asian), 
             num_am_ind = sum(num_am_ind), 
             num_black = sum(num_black), 
             num_hisp = sum(num_hisp))

county_data = county_year_data |>
  group_by(county) |>
  summarize (percent_overweight = mean(percent_overweight), 
             num_white = mean(num_white), 
             num_asian = mean(num_asian), 
             num_am_ind = mean(num_am_ind), 
             num_black = mean(num_black), 
             num_hisp = mean(num_hisp))

county_data = county_data |>
  mutate(total_students = num_am_ind+num_white+num_asian+num_black+num_hisp, 
         percent_asian = num_asian/total_students*100, 
         percent_white = num_white/total_students*100, 
         percent_hisp = num_hisp /total_students*100, 
         percent_am_ind = num_am_ind/total_students*100, 
         percent_black = num_black/total_students*100)
```


```{r message = FALSE, echo = FALSE}
ggplot(county_data, aes(x = county)) +
  geom_bar(aes(y = percent_asian, fill = "Asian"), stat = "identity", position = "stack") +
  geom_bar(aes(y = percent_white, fill = "White"), stat = "identity", position = "stack") +
  geom_bar(aes(y = percent_black, fill = "Black"), stat = "identity", position = "stack") +
  geom_bar(aes(y = percent_am_ind, fill = "Indian"), stat = "identity", position = "stack") +
  geom_bar(aes(y = percent_hisp, fill = "Hispanic"), stat = "identity", position = "stack") +
  geom_line(aes(y = percent_overweight/20, color = "Overweight"), size = 1.5, group = 1) +
  labs(title = "Racial Distribution and Percentage Overweight of Students by County",
       x = "County",
       y = "Percentage of Students",
       fill = "Race") +
  scale_fill_manual(values = c("Asian" = "pink", "White" = "blue", "Black" = "green", "Hispanic" = "purple", "Indian" = "orange")) +
  scale_y_continuous(name = "Percent of different racial students",
                     sec.axis = sec_axis(~.*20, name = "Percentage Overweight")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))


```
interpretation


```{r message = FALSE, echo = FALSE}
# the plotly version of the previous plot
plot_ly(county_data) |>
  add_trace(type = "bar", x = ~county, y = ~percent_asian, name = "Asian", color = "pink") |>
  add_trace(type = "bar", x = ~county, y = ~percent_white, name = "White", color = "blue") |>
  add_trace(type = "bar", x = ~county, y = ~percent_black, name = "Black", color = "green") |>
  add_trace(type = "bar", x = ~county, y = ~percent_am_ind, name = "Indian",color = "orange") |>
  add_trace(type = "bar", x = ~county, y = ~percent_hisp, name = "Hispanic",color = "purple") |>
  add_trace(type = "scatter", x = ~county, y = ~percent_overweight*5, mode = "lines", name = "Overweight", line = list(color = "red", width = 3)) |>
  layout(title = "Racial Distribution and Percentage Overweight of Students by County",
         xaxis = list(title = "County"),
         yaxis = list(title = "Percentage of Students", side = "left"),
         yaxis2 = list(title = "Percentage Overweight", side = "right", overlaying = "y", zeroline = FALSE),
         barmode = "stack",
         showlegend = TRUE)

```

